---
title: "ASSIGNMENT 7.1"
author: "Blaine Blasdell"
date: '2020-07-19'
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: bibliography.bib
---

<!-- ## Library Calls -->
```{r, include=FALSE}
library(rmarkdown)
library(tinytex)
library(boot)
library(ggm)
library(ggplot2)
library(Hmisc)
library(polycor)
library(Rcmdr)
library(readxl)
library(QuantPsyc)
library(foreign)
```
-----------------------------
# Assignment 7.1

<!--Attribute Information:-->

<!-- 1. DGN: Diagnosis - specific combination of ICD-10 codes for primary and secondary as well multiple tumours if any (DGN3,DGN2,DGN4,DGN6,DGN5,DGN8,DGN1) -->
<!-- 2. PRE4: Forced vital capacity - FVC (numeric) -->
<!-- 3. PRE5: Volume that has been exhaled at the end of the first second of forced expiration - FEV1 (numeric) -->
<!-- 4. PRE6: Performance status - Zubrod scale (PRZ2,PRZ1,PRZ0) -->
<!-- 5. PRE7: Pain before surgery (T,F) -->
<!-- 6. PRE8: Haemoptysis before surgery (T,F) -->
<!-- 7. PRE9: Dyspnoea before surgery (T,F) -->
<!-- 8. PRE10: Cough before surgery (T,F) -->
<!-- 9. PRE11: Weakness before surgery (T,F) -->
<!-- 10. PRE14: T in clinical TNM - size of the original tumour, from OC11 (smallest) to OC14 (largest) (OC11,OC14,OC12,OC13) -->
<!-- 11. PRE17: Type 2 DM - diabetes mellitus (T,F) -->
<!-- 12. PRE19: MI up to 6 months (T,F) -->
<!-- 13. PRE25: PAD - peripheral arterial diseases (T,F) -->
<!-- 14. PRE30: Smoking (T,F) -->
<!-- 15. PRE32: Asthma (T,F) -->
<!-- 16. AGE: Age at surgery (numeric) -->
<!-- 17. Risk1Y: 1 year survival period - (T)rue value if died (T,F) -->

```{r, echo=FALSE}
# Read the file ThoraricSurgery.ARFF using read.arff
setwd("D:/Blaine Documents/Bellevue/DSC520/dsc520/Weekly and Projects/data")
thor_df <- read.arff("ThoraricSurgery.ARFF")

```

## Question 7.1a
a. Fit a binary logistic regression model to the data set that predicts whether or not the patient survived for one year (the Risk1Y variable) after the surgery. Use the glm() function to perform the logistic regression. See Generalized Linear Models for an example. Include a summary using the summary() function in your results.

Answer - See output.


```{r, echo=FALSE}
# GLM on original and multi model including summary
thor_glm.1 <- glm(thor_df$Risk1Y ~ DGN, data = thor_df, family = binomial() )
thor_glm.2 <- glm(thor_df$Risk1Y ~ DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30 + PRE32 + AGE, data = thor_df, family = binomial() )
# Test to see if I can remove low significance and raise my accuracy by removing predictors
thor_glm.3 <- glm(thor_df$Risk1Y ~ PRE4 + PRE5 + PRE6 + PRE7  + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE30, data = thor_df, family = binomial() )

summary(thor_glm.1)
summary(thor_glm.2)
summary(thor_glm.3)
```


## Question 7.1b
b. According to the summary, which variables had the greatest effect on the survival rate?

Answer - PRE9 (Dyspnoea before surgery) with a value of True, PRE14 (size of the original tumor) with a diagnosis of OC14 (the largest), PRE17 (type 2 diabetes) with  a diagnosis of True, and PRE30 (Smoking) with a value of true.  


## Question 7.1c
c. To compute the accuracy of your model, use the dataset to predict the outcome variable. The percent of correct predictions is the accuracy of your model. What is the accuracy of your model?

Answer - The accuracy of my initial model with all variables is 83.6 %. By playing with the model I took several less significant variables, and surprisingly including DGN and Age and I got my model to be 84.3% accurate.

```{r, echo=FALSE}

mod1_predict <- predict(thor_glm.1, thor_df, type = "response")
mod2_predict <- predict(thor_glm.2, thor_df, type = "response")
mod3_predict <- predict(thor_glm.3, thor_df, type = "response")

#Confusion Matrix - Validate
conf_max1 <- table(Actual_Value = thor_df$Risk1Yr, Predicted_Value = mod1_predict > 0.5)
conf_max2 <- table(Actual_Value = thor_df$Risk1Yr, Predicted_Value = mod2_predict > 0.5)
conf_max3 <- table(Actual_Value = thor_df$Risk1Yr, Predicted_Value = mod3_predict > 0.5)

conf_max1
conf_max2
conf_max3

# Accuracy
# All Predictors
(conf_max2[[1,1]] + conf_max2[[2,2]]) / sum(conf_max2)
# Predictors removed
(conf_max3[[1,1]] + conf_max3[[2,2]]) / sum(conf_max3)


#

```


# References